Nail: A practical interface generator for data formats

### 2014-05-01 paper/osdi14/eval.tex
pass over the DNS text in 5.1; fix the DNS grammar

To answer the first question, we used Nail to implement grammars
for seven protocols with a range of challenging features.
Figure~\ref{fig:eval-protocols} summarizes these protocols, the lines
of code for their Nail grammars, and the challenging features that make
the protocols difficult to parse with state-of-the-art parser generators.
We find that despite the challenging aspects of these protocols, Nail is
able to capture the protocols, by relying on its novel features: dependent
fields, streams, and transforms.  In contrast, state-of-the-art parser
generators would be unable to fully handle 5 out of the 7 data formats.

@@

\subsubsection{DNS.}

@@

@@

Domain Name System(DNS) packets are described in RFC 1035 and our grammar almost directly
corresponds to the diagrams in Section 4 of that document. Each DNS packet consists of a header, a
set of question records and a set of answer records. Domain names in both queries and answers are
encoded as a sequence of Pascal-style strings. Each label is preceded by a length field, and the
domain name is terminated by a $0$ octet. The client's query packet contains one question
record to the server for each DNS look-up it wants to perform. In addition to the appropriate
response records, the server's reply contains a duplicate of the client's question section, which
allows stateless DNS clients and caches.

To reduce the size overhead of including each domain name twice in a DNS reply, once in the
question section, at least once in the response record, DNS includes a label compression scheme. If
a domain name suffix is repeated, instead of repeating that suffix, the DNS packet may write a
two-bit marker sequence followed by a 14-bit offset into the packet where that suffix was previously
encoded.

Processing these encodings in existing tools, such as Bison or Hammer, would at best be very
awkward, because some ad-hoc trick would have to be used to re-position the parsers input stream.
Keeping track of the position all recognized labels would not be enough, as the offset field may
refer to any byte within the packet, not just the beginning of labels. For this reason, the DNS
server used as a tutorial example for Hammer does not support compression.

Nail handles compression with a stream transform. When parsing, this transform decompresses the DNS
label stream by following the offset pointers. When generating, this transform receives the current
suffix as an input and can scan the packet so far for previous occurrences.

\subsubsection{ZIP Files.}
\input{fig-zip}
\label{s:eval-format-zip}
Another particularly tricky file format is the ZIP compressed archive format\cite{pkzip}.
ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an \emph{end-of-directory
header}. This header contains a variable length comment, so it has to be located by scanning
backwards from the end of the file until a magic number and a valid length field is found. Many ZIP
implementations disagree on how to find this header in confusing situations, such as when the
comment contains the valid magic bytes\cite{wolf:berlinsides-zip}.
This end-of-directory header contains the offset and size of the \emph{ZIP directory}, which is an
array of one \emph{directory entry header} per file in the archive.
Each entry stores file meta data, such as file name, compressed and uncompressed size and a checksum,
in addition to the offset of a \emph{local file header}. The local file header duplicates most
information from the directory entry header and the compressed file contents follow it immediately.

>>>>

To answer the first question, we used Nail to implement grammars
for seven protocols with a range of challenging features.
Figure~\ref{fig:eval-protocols} summarizes these protocols, the lines
of code for their Nail grammars, and the challenging features that make
the protocols difficult to parse with state-of-the-art parser generators.
We find that despite the challenging aspects of these protocols, Nail is
able to capture the protocols, by relying on its novel features: dependent
fields, streams, and transforms.  In contrast, state-of-the-art parser
generators would be unable to fully handle 5 out of the 7 data formats.
In the rest of this subsection, we describe the DNS and Zip grammars in
more detail, focusing on how Nail's features enable us to support these
formats.

@@

\paragraph{DNS.}

Figure~\ref{fig:dns-xform} presents the Nail grammar for DNS packets.
The grammar corresponds almost directly to the diagrams in RFC 1035,
which defines DNS~\cite[\S4]{RFC:1035}.  Each DNS packet consists of a
header, a set of question records, and a set of answer records. Domain
names in both queries and answers are encoded as a sequence of labels,
terminated by a zero byte.  Labels are Pascal-style strings, consisting
of a length field followed by that many bytes comprising the label.

One challenging aspect of DNS packets lies in the count fields (\cc{qc},
\cc{ac}, \cc{ns}, and \cc{ar}), which represent the number of questions
or answers in another part of the packet.  Nail's \cc{n_of} combinator
handles this situation easily, which would have been difficult to handle
for other parsers.

Another challenging aspect of DNS is label
compression~\cite[\S4.1.4]{RFC:1035}.  Label compression is used to reduce
the size overhead of including each domain name multiple times in a DNS
reply (once in the question section, and at least once in the response
section).  If a domain name suffix is repeated, instead of repeating that
suffix, the DNS packet may write a two-bit marker sequence followed by
a 14-bit offset into the packet, indicating the position of where that
suffix was previously encoded.

@@

Handling label compression in existing tools, such as Bison or Hammer,
would at best be very awkward, because some ad-hoc trick would have to
be used to re-position the parser's input stream.  Keeping track of the
position of all recognized labels would not be enough, as the offset field
may refer to any byte within the packet, not just the beginning of labels.
For this reason, the DNS server used as the example for Hammer does not
support compression.

@@

In contrast, Nail is able to handle label compression, by using a stream
transform; the signatures of the two transform functions are shown
in Figure~\ref{fig:dns-xform}.  When parsing a packet, this transform
decompresses the DNS label stream by following the offset pointers.
When generating a packet, this transform receives the current suffix as
an input, and scans the packet so far for previous occurrences, which
implements compression.

\paragraph{ZIP files.}

\input{fig-zip}
\label{s:eval-format-zip}
Another particularly tricky file format is the ZIP compressed archive format\cite{pkzip}.
ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an \emph{end-of-directory
header}. This header contains a variable length comment, so it has to be located by scanning
backwards from the end of the file until a magic number and a valid length field is found. Many ZIP
implementations disagree on how to find this header in confusing situations, such as when the
comment contains the valid magic bytes\cite{wolf:berlinsides-zip}.
This end-of-directory header contains the offset and size of the \emph{ZIP directory}, which is an
array of one \emph{directory entry header} per file in the archive.
Each entry stores file meta data, such as file name, compressed and uncompressed size and a checksum,
in addition to the offset of a \emph{local file header}. The local file header duplicates most
information from the directory entry header and the compressed file contents follow it immediately.
